Higher-Order unification for free!

### 2024-04-12 paper/main.tex
sec 2.2

We write $\sigma = \{~ X \mapsto t ~\}$ for the substitution that assigns
the term $t$ to the variable $X$.
We write $\sigma t$ for the application of
the substitution to a term $t$, and $\sigma X = \{~ \sigma t ~|~ t \in X ~\}$ when
$X$ is a set of terms.
We write $\sigma \subseteq \sigma'$ when $\sigma$ is more
general than $\sigma'$.
The domain of a substitution is the set of unification variables for which
it provides an assignment.
We write $\sigma \cup \sigma'$ to denote the concatenation of
two substitutions whose domains are disjoint.
We shall use $\rho$ for \Fo{} substitutions,
and $\sigma$ for the \Ho ones.
For brevity, in this section we consider
the substitution for \Fo{} and \Ho{} identical.
We defer to \cref{sec:grounwork} a more precise description
pointing out theirs differences.

@@

We write \todo{meglio dire che sigma' estende sigma, togliere unione e quindi dominio da sopra}
$\sigma t_1 \Ue \sigma t_2 \mapsto \sigma'$ when
$\sigma t_1$ and $\sigma t_2$ unify with substitution $\sigma'$.
We write $t_1 \Ue t_2 \mapsto \sigma$ when
the initial substitution is empty.
Note that if $\sigma t_1 \Ue \sigma t_2 \mapsto \sigma'$ then
the domains of $\sigma$ and $\sigma'$ are disjoint.

@@

\todo{spiegare llam (first use after intro)}
Even if we provide an implementation of the object-language unification
\Uo{} in \cref{sec:founif}, our real goal is the simulation of an entire
logic program.

@@

\begin{lemma}[\wellb-enforcement]\label{lem:w-enforcement-maybebeta} Even if $\foUnifPb \cap \maybebeta \ne \emptyset$,
  $\hoUnifPb \cup \maybebeta = \emptyset$
\end{lemma}
\begin{proof}[Proof sketch]
  problematic terms are mapped to uva by comp, the problematic fapp node is gone.
\end{proof}

>>>>

We write $\sigma = \{~ X \mapsto t ~\}$ for the substitution that assigns
the term $t$ to the variable $X$.
We write $\sigma t$ for the application of
the substitution to a term $t$, and $\sigma X = \{~ \sigma t ~|~ t \in X ~\}$ when
$X$ is a set of terms.
We write $\sigma \subseteq \sigma'$ when $\sigma$ is more
general than $\sigma'$.
% The domain of a substitution is the set of unification variables for which
% it provides an assignment.
% We write $\sigma \cup \sigma'$ to denote the concatenation of
% two substitutions whose domains are disjoint.
We shall use $\rho$ for \Fo{} substitutions,
and $\sigma$ for the \Ho ones.
For brevity, in this section we consider
the substitution for \Fo{} and \Ho{} identical.
We defer to \cref{sec:grounwork} a more precise description
pointing out theirs differences.

@@

We write 
$\sigma t_1 \Ue \sigma t_2 \mapsto \sigma'$ when
$\sigma t_1$ and $\sigma t_2$ unify with substitution $\sigma'$.
Note that $\sigma'$ is a refined (i.e. extended) version of $\sigma$.
We write $t_1 \Ue t_2 \mapsto \sigma'$ when
the initial substitution $\sigma$ is empty.
% Note that if $\sigma t_1 \Ue \sigma t_2 \mapsto \sigma'$ then
% the domains of $\sigma$ and $\sigma'$ are disjoint.
We write \llambda as the set of terms that are in the pattern-fragment, i.e.
every higher-order variable is applied to a list of distinct names.

@@

Even if we provide an implementation of the object-language unification
\Uo{} in \cref{sec:founif}, our real goal is the simulation of an entire
logic program.

@@

\begin{lemma}[\wellb-enforcement]\label{lem:w-enforcement-maybebeta} Even if $\foUnifPb \cap \maybebeta \ne \emptyset$,
  $\hoUnifPb \cup \maybebeta = \emptyset$
\end{lemma}\noteD{Non capisco}
\begin{proof}[Proof sketch]
  problematic terms are mapped to uva by comp, the problematic fapp node is gone.
\end{proof}
